---
title: "Question 2"
output: html_document
date: "2022-12-16"
---

```{r}
# Initializing parameters
set.seed(320733099)

# Generating the random variables
rg <- function(n){
  t = runif(n)
  sin(t)*exp(-3*t^2)/(1+t)
}

N=10000
k=0.01
alpha=0.05
n=500
```

The goal is to find the minimum value $n \in \mathbb{N}$ such that $\mathbb{P}(|\hat{I}_n - I| > 0.01) \leq 0.05$. In order to do this, we can observe that since $\hat{I}_n$ is unbiased for $I$, $$\frac{\hat{I}_n - I}{\left(\frac{S}{\sqrt{n}}\right)} \sim \mathcal{N}(0,1)$$ by the Central Limit Theorem with $S$ being the sample standard deviation. So then we have $$\begin{align}\mathbb{P}(|\hat{I}_n - I| > 0.01) &\leq 0.05\\
1-\mathbb{P}(-0.01 < \hat{I}_n - I < 0.01) &\leq 0.05\\
1-\mathbb{P}\left(\frac{-0.01}{\frac{S}{\sqrt{n}}} < \frac{\hat{I}_n - I}{\frac{S}{\sqrt{n}}} < \frac{0.01}{\frac{S}{\sqrt{n}}}\right) &\leq 0.05\\
\tag{Z is a standard normal}1-\mathbb{P}\left(\frac{-0.01\sqrt{n})}{S} < Z < \frac{0.01\sqrt{n}}{S}\right) &\leq 0.05\\
2*\mathbb{P}\left( Z > \frac{0.01\sqrt{n}}{S}\right) &\leq 0.05\\
\mathbb{P}\left( Z > \frac{0.01\sqrt{n}}{S}\right) &\leq 0.025\\
\mathbb{P}\left( Z < \frac{0.01\sqrt{n}}{S}\right) &\leq 0.975\\
\frac{0.01\sqrt{n}}{S} &\geq \Phi^{-1}(0.975)\\
n &\geq \left(\frac{S\cdot \Phi^{-1}(0.975)}{0.01}\right)^2
\end{align}$$ where we can double since $Z$ is symmetric. First, we will do a simulation by running the $I$ distribution with $n=500$ samples and we'll obtain the sample variance. Then we will use that to calculate the best estimate for $n$ given the equation above. We will take the ceiling of whatever the estimate comes out to be. One issue is that we do not know what the true mean for $I$ is. However, we do know that $\frac{1}{n}\sum_{i=1}^n I_i \xrightarrow{p} E[I]$ by the weak law of large numbers. Therefore, we can let the sample mean be our estimate for $E[I]$.

```{r}
samp = rg(n)
mu_est = mean(samp)
sig_est = sd(samp)
n_est = ceiling((sig_est*qnorm(1-alpha/2)/k)^2)

data.frame(sample_mean=mu_est,sample_sd=sig_est, n=n_est)
```
After taking $500$ samples, we get an estimate of $\hat{I}_n\approx 0.108$ and $s\approx 0.0538$. Then we compute the most suitable $n$ to be $n \approx 112$. Now we can run $N=10000$ repetitions of sample sizes $112$ to verify if $\mathbb{P}(|\hat{I}_n - I| > 0.01) \leq 0.05$. We know that $E[\hat_{I}_n]=I$. However, we do not know what the true $I$ is. Therefore, we can approximate it with the sample mean. It is worthy of note that using the sample mean will not be perfectly representative. If we use the estimate above `mu_est`, we expect to get an over-estimate because of the variance in $\hat{I}_n$. It will be slightly off of $I$ which leaves more room for the difference to be greater than $0.01$. On the other hand, if we compute the sample mean for the data we are running the tests on `mean_of_samples`, we expect to get an underestimate. This is due to the fact that the sample mean has the least variance from the sampled points.

Taking all of this into account, I'd say this calculation is a success if `mu_est` is slightly greater than $0.05$ and if `mean_of_samples` is slightly less than $0.05$.

```{r}
# running the samples
samp_means = rep(0,N)
for(i in 1:N){
  I_n = rg(n_est)
  samp_means[i] = mean(I_n)
}
# taking mean of sample means to get an approximate value for the true mean.
mean_of_samples = mean(samp_means)

# show the proportion with two different estimates.
predetermined=length(which(abs(samp_means - mu_est) > k))/N
from_samples=length(which(abs(samp_means - mean_of_samples) > k))/N

data.frame(prior_estimate=predetermined, mean_of_samples=from_samples)
```

By comparing to the sample mean of a separate sample `mu_est`, we get that a proportion of $0.0742$ samples outside the range. By comparing to the sample mean of the test samples, we get a proportion of $0.0447$ samples outside the range. These results are exactly what I expected with the way the sample means are calculated.