---
title: "Question 7"
output: html_document
date: "2022-12-16"
---

```{r}
set.seed(1105677292)
library(MASS)
#Initializing parameters
B=500
N=5000
mu=c(8,4,1,2)
sig=matrix(c(6,2,1,0,
             2,6,1,1,
             1,1,3,0,
             0,1,0,7)
           ,nrow = 4,ncol=4)
```

For this algorithm we will separate the $X_1$ and $X_2$ from $X_3$ and $X_4$. This way, we can use the conditional distribution to use a bivariate normal. Let $$Y_1=\begin{bmatrix}X_1\\
X_2\end{bmatrix}$$ and $$Y_2=\begin{bmatrix}X_3\\
X_4\end{bmatrix}$$ We will use the fact that $$Y_1 | Y_2 \sim N(\hat{y}_1, \Sigma_{(11)}-\Sigma_{(12)}\Sigma_{(22)}^{-1}\Sigma_{(21)})$$ where $$\hat{y}_1=\begin{bmatrix}\mu_1\\
\mu_2\end{bmatrix} + \Sigma_{(12)}\Sigma_{(22)}^{-1}(Y_2 - \begin{bmatrix}\mu_3\\
\mu_4\end{bmatrix})$$ and $\Sigma_{ij}$ corresponds to the $i$th vertically, $j$th horizontally submatrix. So for example, $$\Sigma_{11}=\begin{bmatrix}6 & 2\\
2 & 6\end{bmatrix}.$$ Using this, we can compute the conditional distributions to find the posterior distributions.

```{r}
#initialize by drawing from x_i each independently
sample_data=matrix(0,nrow = B+N,ncol = 4)
sample_data[1,]=rnorm(4,mean=mu,sd=sqrt(diag(sig)))
sub11=sig[1:2,1:2]
sub12=sig[1:2,3:4]
sub21=sig[3:4,1:2]
sub22=sig[3:4,3:4]

for(i in 2:(B+N)){
  
#conditional distributions
sample_data[i,1:2]=mvrnorm(n = 1,
                               mu = mu[1:2]+sub12%*%solve(sub22)%*%(sample_data[i-1,3:4]-mu[3:4]), 
                               Sigma = sub11-sub12%*%solve(sub22)%*%sub21)

sample_data[i,3:4]=mvrnorm(n = 1,
                               mu = mu[3:4]+sub21%*%solve(sub11)%*%(sample_data[i,1:2]-mu[1:2]), 
                               Sigma = sub22-sub21%*%solve(sub11)%*%sub12)
}

#take out the burn in samples
data=sample_data[B+1:N,]
expectation=mean(log(abs(data[,1]/data[,3]))*abs(data[,2]/data[,4]))
print(expectation)
```

After running the gibbs algorithm and taking the sample mean, we can approximate that $$E[\log(|X_1/X_3|)\times |X_2/X_4|]\approx 14.56$$ After running this a few times though, I notice that there is a TON of variance. Even with a sample size of 5000, the sample mean values range from $11-366$.My guess is that when $X_1$ or $X_4$ get extremely close to 0, the value blows up into something extremely large which causes this massive range of values. It surely isn't normally distributed.
